---
title: "R Notebook"
output: html_notebook
---

```{r global_options}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
```

```{r}
library(tidyverse)
library(broom)
library(readxl)
APLS_e <- read_excel("C:/Users/Supun/Desktop/APLS/analysis/APLS/APLS_e.xlsx")
```


```{r}
full <- APLS_e
```

Total mothers
```{r}
print(length(full$`Serial Number`))
```

```{r}

asp <- full %>% pivot_longer(cols = str_subset(names(full), "_Aspirin"), names_to = "pregnancy", values_to = "ASP")

lm <- full %>% pivot_longer(cols = str_subset(names(full), "_LMWH"), names_to = "pregnancy", values_to = "LMWH") %>% select(LMWH)

out <- full %>% pivot_longer(cols = str_subset(names(full), "_outcome"), names_to = "pregnancy_", values_to = "outcome") %>% select(outcome)

pih <- full %>% pivot_longer(cols = str_subset(names(full), "_PIH"), names_to = "pregnancy", values_to = "PIH") %>% select(PIH)

poa <- full %>% select(str_subset(names(full), "_POA")) %>% mutate_if(is.character, as.double) %>% pivot_longer(cols = str_subset(names(full), "_POA"), names_to = "pregnancy", values_to = "POA") %>% select(POA)


com <- cbind(asp,lm, out, pih, poa) %>% select(- str_subset(names(asp), "_LMWH")) %>% select(- str_subset(names(asp), "_outcome")) %>% select(- str_subset(names(asp), "_PIH")) %>% select(- str_subset(names(asp), "_POA")) %>% filter(!is.na(outcome)) %>% mutate(pregnancy = str_extract(pregnancy, "\\d{1,2}")) 

com %>% summarise(total_pregnancies = n())

com %>%mutate(LMWH = ifelse(is.na(LMWH),0,LMWH), ASP = ifelse(is.na(ASP),0,ASP)) %>% summarize(Aspirin = sum(ASP), LMWH = sum(LMWH))

com %>%mutate(LMWH = ifelse(is.na(LMWH),0,LMWH), ASP = ifelse(is.na(ASP),0,ASP)) %>% summarize(Both = sum(LMWH == 1 & ASP == 1))

com %>% mutate(LMWH = ifelse(is.na(LMWH),0,LMWH), ASP = ifelse(is.na(ASP),0,ASP)) %>%
summarize(None = sum(LMWH == 0 & ASP == 0))
```

demography

number of pregnancies 
```{r}
com %>% group_by(`Serial Number`) %>% summarise(n = n()) %>% pull(n) %>% quantile()
com %>% group_by(`Serial Number`) %>% summarise(n = n()) %>% ggplot(aes(n)) + geom_boxplot()
```

Age
```{r}
com %>% group_by(`Serial Number`) %>% summarise(age = min(Age)) %>% summarise(mean = mean(age, na.rm = TRUE), sd = sd(age, na.rm = TRUE))

com %>% group_by(`Serial Number`) %>% summarise(age = min(Age)) %>% ggplot(aes(age)) + geom_histogram(binwidth = 2)


```

PIH
```{r}
com %>% summarise(PIH = sum(PIH, na.rm = TRUE), "percent - out of pregnancies" = sum(PIH, na.rm = TRUE) * 100 / n() )
```

Other Medical conditions
```{r}
com %>% filter(!is.na(`Medical Conditions`)) %>% group_by(`Serial Number`) %>% summarise(condition = min(`Medical Conditions`)) %>% group_by(condition) %>% summarise(n = n(), "percent - out of mothers" = n() * 100 / length(full$`Serial Number`))
```

Investigations
Any of the 5 investigations
```{r}
com %>% group_by(`Serial Number`) %>% 
  summarise(any = ifelse(!is.na(min(Lupus)) | !is.na(min(ANA)) | !is.na(min(Cardiolipin)) | !is.na(min(APLS)) | !is.na(min(`beta2 microglobulin`)), 1,0 )) %>% 
  summarise("Any investigation" = sum(any), "precent - of mothers" = sum(any)*100/n())
```

Any of  Lupus, APLS, cardiolipin, beta2
```{r}
com %>% group_by(`Serial Number`) %>% 
  summarise(any = ifelse(!is.na(min(Lupus)) | !is.na(min(Cardiolipin)) | !is.na(min(APLS)) | !is.na(min(`beta2 microglobulin`)), 1,0 )) %>% 
  summarise("Any investigation out of Lupus, cardio, beta, APLS" = sum(any), "precent - of mothers" = sum(any)*100/n())
```

Proportion of a positive result

```{r}
com %>% group_by(`Serial Number`) %>% 
  summarise(any = ifelse(!is.na(min(Lupus)) | !is.na(min(Cardiolipin)) | !is.na(min(APLS)) | !is.na(min(`beta2 microglobulin`)), 1,0 ), any_positive = ifelse(min(Lupus) == 1 | min(Cardiolipin) == 1  | min(APLS) == 1 | min(`beta2 microglobulin`) == 1, 1, 0)) %>% 
  mutate(any_positive = ifelse(is.na(any_positive),0, any_positive)) %>% 
  filter(any == 1) %>%
  summarise("positive subjects" = sum(any_positive), "percent of tested" = sum(any_positive) * 100 / n())
```


Checking for clinical criteria

At or beyond 10th wk
```{r}
com %>% 
  mutate(live = ifelse(outcome %in% c("EL-LSCS", "EM-LSCS", "Live", "LIVE", "LSCS", "Neonatal Death"),1,0)) %>% 
  filter(live == 0) %>%
  filter(POA >= 10) %>% group_by(`Serial Number`) %>% 
  summarise(n = n()) %>% summarise("MC at or over 10wk" = n())
```

3 or more MC before 10 wk
```{r}
com %>% 
    mutate(live = ifelse(outcome %in% c("EL-LSCS", "EM-LSCS", "Live", "LIVE", "LSCS", "Neonatal Death"),1,0)) %>% 
    filter(live == 0) %>%
    filter(POA < 10) %>% group_by(`Serial Number`) %>% 
    filter(n() >= 3) %>%
    summarise(n = n()) %>% summarise("at least 3 MC before 10 wk" = n())
```

Premature births before 34 wk
```{r}
com %>% 
    mutate(live = ifelse(outcome %in% c("EL-LSCS", "EM-LSCS", "Live", "LIVE", "LSCS", "Neonatal Death"),1,0)) %>% 
    filter(live == 1) %>%
    filter(POA < 34) %>% group_by(`Serial Number`) %>% 
    summarise(n = n()) %>% summarise("Premature birth before 34 wk" = n())
```


Outcome
```{r}
com <- com %>% 
    mutate(live = ifelse(outcome %in% c("EL-LSCS", "EM-LSCS", "Live", "LIVE", "LSCS", "Neonatal Death"),1,0))

```


Previous pregnancy outcome without specific
treatment
```{r}
com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 0 & LMWH == 0) %>% summarise('total pregnancies' = n())

com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 0 & LMWH == 0) %>% filter(live == 1) %>% summarise('successful pregnancies' = n())

no_treatment_total <- com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 0 & LMWH == 0) %>% summarise('total pregnancies' = n()) %>% pull(`total pregnancies`)

no_treatment_success <- com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 0 & LMWH == 0) %>% filter(live == 1) %>% summarise('successful pregnancies' = n()) %>% pull('successful pregnancies')

com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 0 & LMWH == 0) %>% filter(live == 0) %>% summarise('failed pregnancies' = n())

com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 0 & LMWH == 0) %>% summarise('success rate' = mean(live)*100)

com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 0 & LMWH == 0) %>% filter(live == 0 & POA < 10) %>% summarise('MC before 10 wk' = n())

com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 0 & LMWH == 0) %>% filter(live == 0 & POA >= 10) %>% summarise('MC at or after 10 wk' = n())
```


Previous pregnancy outcome - Treated by either Aspirin or LMWH
```{r}
com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 1 | LMWH == 1) %>% summarise('total pregnancies' = n())

com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 1 | LMWH == 1) %>% filter(live == 1) %>% summarise('successful pregnancies' = n())

either_treatment_total <- com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 1 | LMWH == 1) %>% summarise('total pregnancies' = n()) %>% pull(`total pregnancies`)

either_treatment_success <- com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 1 | LMWH == 1) %>% filter(live == 1) %>% summarise('successful pregnancies' = n()) %>% pull('successful pregnancies')

com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 1 | LMWH == 1) %>% filter(live == 0) %>% summarise('failed pregnancies' = n())

com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 1 | LMWH == 1) %>% summarise('success rate' = mean(live)*100)

com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 1 | LMWH == 1) %>% filter(live == 0 & POA < 10) %>% summarise('MC before 10 wk' = n())

com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 1 | LMWH == 1) %>% filter(live == 0 & POA >= 10) %>% summarise('MC at or after 10 wk' = n())
```


Previous pregnancy outcome - Treated Aspirin and LMWH
```{r}
com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 1 & LMWH == 1) %>% summarise('total pregnancies' = n())

com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 1 & LMWH == 1) %>% filter(live == 1) %>% summarise('successful pregnancies' = n())

both_treatment_total <- com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 1 & LMWH == 1) %>% summarise('total pregnancies' = n()) %>% pull(`total pregnancies`)

both_treatment_success <- com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 1 & LMWH == 1) %>% filter(live == 1) %>% summarise('successful pregnancies' = n()) %>% pull('successful pregnancies')

com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 1 & LMWH == 1) %>% filter(live == 0) %>% summarise('failed pregnancies' = n())

com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 1 & LMWH == 1) %>% summarise('success rate' = mean(live)*100)

com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 1 & LMWH == 1) %>% filter(live == 0 & POA < 10) %>% summarise('MC before 10 wk' = n())

com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 1 & LMWH == 1) %>% filter(live == 0 & POA >= 10) %>% summarise('MC at or after 10 wk' = n())
```


Previous pregnancy outcome - Treated Aspirin alone
```{r}
com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 1 & LMWH == 0) %>% summarise('total pregnancies' = n())

com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 1 & LMWH == 0) %>% filter(live == 1) %>% summarise('successful pregnancies' = n())

asp_treatment_total <- com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 1 & LMWH == 0) %>% summarise('total pregnancies' = n()) %>% pull(`total pregnancies`)

asp_treatment_success <- com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 1 & LMWH == 0) %>% filter(live == 1) %>% summarise('successful pregnancies' = n()) %>% pull('successful pregnancies')

com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 1 & LMWH == 0) %>% filter(live == 0) %>% summarise('failed pregnancies' = n())

com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 1 & LMWH == 0) %>% summarise('success rate' = mean(live)*100)

com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 1 & LMWH == 0) %>% filter(live == 0 & POA < 10) %>% summarise('MC before 10 wk' = n())

com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 1 & LMWH == 0) %>% filter(live == 0 & POA >= 10) %>% summarise('MC at or after 10 wk' = n())
```


Previous pregnancy outcome - Treated LMWH alone
```{r}
com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 0 & LMWH == 1) %>% summarise('total pregnancies' = n())

com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 0 & LMWH == 1) %>% filter(live == 1) %>% summarise('successful pregnancies' = n())


LMWH_treatment_total <- com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 0 & LMWH == 1) %>% summarise('total pregnancies' = n()) %>% pull(`total pregnancies`)

LMWH_treatment_success <- com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 0 & LMWH == 1) %>% filter(live == 1) %>% summarise('successful pregnancies' = n()) %>% pull('successful pregnancies')

com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 0 & LMWH == 1) %>% filter(live == 0) %>% summarise('failed pregnancies' = n())

com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 0 & LMWH == 1) %>% summarise('success rate' = mean(live)*100)

com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 0 & LMWH == 1) %>% filter(live == 0 & POA < 10) %>% summarise('MC before 10 wk' = n())

com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 0 & LMWH == 1) %>% filter(live == 0 & POA >= 10) %>% summarise('MC at or after 10 wk' = n())
```


comparison

No treatment vs both
```{r}
f1 <- prop.test(c(both_treatment_success, no_treatment_success), c(both_treatment_total, no_treatment_total), p = NULL, alternative = "greater",correct = TRUE)
print(tidy(f1))
```

No treatment vs asp
```{r}
f1 <- prop.test(c(asp_treatment_success, no_treatment_success), c(asp_treatment_total, no_treatment_total), p = NULL, alternative = "greater",correct = TRUE)
print(tidy(f1))
```


No treatment vs LMWH
```{r}
f1 <- prop.test(c(LMWH_treatment_success, no_treatment_success), c(LMWH_treatment_total, no_treatment_total), p = NULL, alternative = "greater",correct = TRUE)
print(tidy(f1))
```

ASP treatment vs LMWH
```{r}
f1 <- prop.test(c(LMWH_treatment_success, asp_treatment_success), c(LMWH_treatment_total, asp_treatment_total), p = NULL, alternative = "two.sided",correct = TRUE)
print(tidy(f1))
```

ASP treatment vs both
```{r}
f1 <- prop.test(c(both_treatment_success, asp_treatment_success), c(both_treatment_total, asp_treatment_total), p = NULL, alternative = "greater",correct = TRUE)
print(tidy(f1))
```

LMWH treatment vs both
```{r}
f1 <- prop.test(c(both_treatment_success, LMWH_treatment_success), c(both_treatment_total, LMWH_treatment_total), p = NULL, alternative = "greater",correct = TRUE)
print(tidy(f1))
```

Live rates between test positives and negatives

```{r}
p <- com %>%
  mutate(any = ifelse(!is.na(Lupus) | !is.na((Cardiolipin)) | !is.na((APLS)) | !is.na((`beta2 microglobulin`)), 1,0 ), any_positive = ifelse((Lupus) == 1 | (Cardiolipin) == 1  | (APLS) == 1 | (`beta2 microglobulin`) == 1, 1, 0)) %>%
  mutate(any_positive = ifelse(is.na(any_positive),0, any_positive)) %>%
  filter(any == 1) %>% filter(ASP == 1 | LMWH == 1) %>% group_by(any_positive) %>% summarise(n = n(), live_number = sum(live), live_rate = mean(live))
print(p)
print(tidy(prop.test(p$live_number, p$n)))
```

number of pregnancies that were treated, in mothers who were tested for antibodies was `r sum(p$n)`.
There ware `r p$n[2]` mothers tested positive for at least one test and the success rate was `r p$live_rate[2]`.
`r p$n[1]` mothers had negative results and Success rate among them was `r p$live_rate[1]`