---
title: "APLS study"
output: 
  html_notebook: 
    toc: yes
    code_folding: hide
    fig_caption: yes
---


```{r global_options, message=FALSE, warning=FALSE}

knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)

```

```{r}
library(tidyverse)
library(broom)
library(readxl)
APLS_e <- read_excel("APLS_e.xlsx")
```


```{r}
full <- APLS_e
```

Total mothers
```{r}
print(length(full$`Serial Number`))
total_mothers <- length(full$`Serial Number`)
```

```{r}
# pivoting

asp <- full %>% pivot_longer(cols = str_subset(names(full), "_Aspirin"), names_to = "pregnancy", values_to = "ASP")

lm <- full %>% pivot_longer(cols = str_subset(names(full), "_LMWH"), names_to = "pregnancy", values_to = "LMWH") %>% select(LMWH)

out <- full %>% pivot_longer(cols = str_subset(names(full), "_outcome"), names_to = "pregnancy_", values_to = "outcome") %>% select(outcome)

pih <- full %>% pivot_longer(cols = str_subset(names(full), "_PIH"), names_to = "pregnancy", values_to = "PIH") %>% select(PIH)

poa <- full %>% select(str_subset(names(full), "_POA")) %>% mutate_if(is.character, as.double) %>% pivot_longer(cols = str_subset(names(full), "_POA"), names_to = "pregnancy", values_to = "POA") %>% select(POA)


com <- cbind(asp,lm, out, pih, poa) %>% select(- str_subset(names(asp), "_LMWH")) %>% select(- str_subset(names(asp), "_outcome")) %>% select(- str_subset(names(asp), "_PIH")) %>% select(- str_subset(names(asp), "_POA")) %>% filter(!is.na(outcome)) %>% mutate(pregnancy = str_extract(pregnancy, "\\d{1,2}")) 

com %>% summarise(total_pregnancies = n())
total_pregnancies <- com %>% summarise(total_pregnancies = n()) %>% pull(total_pregnancies)

```

Checking for clinical criteria
```{r}
com <- com %>% 
    mutate(live = ifelse(outcome %in% c("EL-LSCS", "EM-LSCS", "Live", "LIVE", "LSCS", "Neonatal Death"),1,0))

is_con <- function(v){
  result <- rle(diff(as.numeric(v)))
  any(result$lengths>=2 & result$values==1)
}

cr_1 <- com %>% filter(POA >= 10, live == 0) %>% group_by(`Serial Number`) %>% summarise(n = n(), criterion = 1)

cr_2 <- com %>% filter(!`Serial Number` %in% cr_1$`Serial Number`) %>% filter(live == 1 & POA < 34) %>% group_by(`Serial Number`) %>% summarise(n = n(), criterion = 2)

cr_3 <- com %>% filter(!`Serial Number` %in% cr_1$`Serial Number` & live == 0) %>% filter(!is.na(POA)) %>% group_by(`Serial Number`) %>% filter(is_con(pregnancy)) %>% summarise(n = n(), criterion = 3)

no_cr <- com %>% filter(!`Serial Number` %in% cr_1$`Serial Number` & !`Serial Number` %in% cr_2$`Serial Number` & !`Serial Number` %in% cr_3$`Serial Number`) %>% group_by(`Serial Number`) %>% summarise(n = n(), criterion = "no")

com_all <- com

#selecting only eligible
com <- com %>% filter(`Serial Number` %in% cr_1$`Serial Number` | `Serial Number` %in% cr_2$`Serial Number`  | `Serial Number` %in% cr_3$`Serial Number`)

rbind(cr_1,cr_2,cr_3,no_cr) %>% group_by(criterion) %>% summarise(n = n())
total_mothers <- com %>% group_by(`Serial Number`) %>% summarise(n = n()) %>% summarise(n = n()) %>% pull(n)
total_pregnancies <- com %>% summarise(n = n()) %>% pull(n)
```


After correcting for criteria

```{r}
knitr::kable(data.frame("Total mothers" = total_mothers, "Total pregnancie" = total_pregnancies))
```
Treatment
```{r}
# treatment

treatment <- com %>% mutate(LMWH = ifelse(is.na(LMWH),0,LMWH), ASP = ifelse(is.na(ASP),0,ASP)) %>% summarize(Aspirin_only = sum(ASP == 1 & LMWH == 0, na.rm = TRUE), LMWH_only = sum(ASP == 0 & LMWH == 1), Either = sum(ASP == 1 | LMWH == 1),None = sum(LMWH == 0 & ASP == 0), Both = sum(LMWH == 1 & ASP == 1), Aspirin = sum(ASP), LMWH = sum(LMWH))

print(treatment)
```

demography

number of pregnancies 
```{r message=FALSE, warning=FALSE}
com %>% group_by(`Serial Number`) %>% summarise(n = n()) %>% pull(n) %>% quantile()

pregnancy_quantiles <- com %>% group_by(`Serial Number`) %>% summarise(n = n()) %>% pull(n) %>% quantile()

pregnancy_boxplot <- com %>% ggplot(aes(as.numeric(pregnancy))) + geom_boxplot()
```

Age
```{r}
age <- com %>% group_by(`Serial Number`) %>% summarise(age = min(Age)) %>% summarise(mean = mean(age, na.rm = TRUE), sd = sd(age, na.rm = TRUE))

age_histogram <- com %>% group_by(`Serial Number`) %>% summarise(age = min(Age)) %>% ggplot() + geom_histogram(aes(x = age), binwidth = 2, fill = "white", col = "black")


```

PIH
```{r}
pih <- com %>% summarise(PIH = sum(PIH, na.rm = TRUE), "percent - out of pregnancies" = sum(PIH, na.rm = TRUE) * 100 / n() )
print(pih)
```

Other Medical conditions
```{r}
com %>% filter(!is.na(`Medical Conditions`)) %>% group_by(`Serial Number`) %>% summarise(condition = min(`Medical Conditions`)) %>% group_by(condition) %>% summarise(n = n(), "percent - out of mothers" = n() * 100 / length(full$`Serial Number`))
```

Investigations
Any of the 5 investigations
```{r}
com %>% group_by(`Serial Number`) %>% 
  summarise(any = ifelse(!is.na(min(Lupus)) | !is.na(min(ANA)) | !is.na(min(Cardiolipin)) | !is.na(min(APLS)) | !is.na(min(`beta2 microglobulin`)), 1,0 )) %>% 
  summarise("Any investigation" = sum(any), "precent - of mothers" = sum(any)*100/n())
```

Any of  Lupus, APLS, cardiolipin, beta2
```{r}
investigations_all <- com %>% group_by(`Serial Number`) %>% 
  summarise(any = ifelse(!is.na(min(Lupus)) | !is.na(min(Cardiolipin)) | !is.na(min(APLS)) | !is.na(min(`beta2 microglobulin`)), 1,0 )) %>% 
  summarise("Any investigation out of Lupus, cardio, beta, APLS" = sum(any), "precent - of mothers" = sum(any)*100/n())
```

Proportion of a positive result

```{r}
investigations_positive <- com %>% group_by(`Serial Number`) %>% 
  summarise(any = ifelse(!is.na(min(Lupus)) | !is.na(min(Cardiolipin)) | !is.na(min(APLS)) | !is.na(min(`beta2 microglobulin`)), 1,0 ), any_positive = ifelse(min(Lupus) == 1 | min(Cardiolipin) == 1  | min(APLS) == 1 | min(`beta2 microglobulin`) == 1, 1, 0)) %>% 
  mutate(any_positive = ifelse(is.na(any_positive),0, any_positive)) %>% 
  filter(any == 1) %>%
  summarise("positive subjects" = sum(any_positive), "percent of tested" = sum(any_positive) * 100 / n())
investigations_all
investigations_positive
```

Previous pregnancy outcome without specific
treatment
```{r}
com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 0 & LMWH == 0) %>% summarise('total pregnancies' = n())

com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 0 & LMWH == 0) %>% filter(live == 1) %>% summarise('successful pregnancies' = n())

no_treatment_total <- com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 0 & LMWH == 0) %>% summarise('total pregnancies' = n()) %>% pull(`total pregnancies`)

no_treatment_success <- com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 0 & LMWH == 0) %>% filter(live == 1) %>% summarise('successful pregnancies' = n()) %>% pull('successful pregnancies')

com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 0 & LMWH == 0) %>% filter(live == 0) %>% summarise('failed pregnancies' = n())

com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 0 & LMWH == 0) %>% summarise('success rate' = mean(live)*100)

com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 0 & LMWH == 0) %>% filter(live == 0 & POA < 10) %>% summarise('MC before 10 wk' = n())

com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 0 & LMWH == 0) %>% filter(live == 0 & POA >= 10) %>% summarise('MC at or after 10 wk' = n())
```


Previous pregnancy outcome - Treated by either Aspirin or LMWH
```{r}
com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 1 | LMWH == 1) %>% summarise('total pregnancies' = n())

com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 1 | LMWH == 1) %>% filter(live == 1) %>% summarise('successful pregnancies' = n())

either_treatment_total <- com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 1 | LMWH == 1) %>% summarise('total pregnancies' = n()) %>% pull(`total pregnancies`)

either_treatment_success <- com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 1 | LMWH == 1) %>% filter(live == 1) %>% summarise('successful pregnancies' = n()) %>% pull('successful pregnancies')

com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 1 | LMWH == 1) %>% filter(live == 0) %>% summarise('failed pregnancies' = n())

com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 1 | LMWH == 1) %>% summarise('success rate' = mean(live)*100)

com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 1 | LMWH == 1) %>% filter(live == 0 & POA < 10) %>% summarise('MC before 10 wk' = n())

com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 1 | LMWH == 1) %>% filter(live == 0 & POA >= 10) %>% summarise('MC at or after 10 wk' = n())
```


Previous pregnancy outcome - Treated Aspirin and LMWH
```{r}
com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 1 & LMWH == 1) %>% summarise('total pregnancies' = n())

com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 1 & LMWH == 1) %>% filter(live == 1) %>% summarise('successful pregnancies' = n())

both_treatment_total <- com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 1 & LMWH == 1) %>% summarise('total pregnancies' = n()) %>% pull(`total pregnancies`)

both_treatment_success <- com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 1 & LMWH == 1) %>% filter(live == 1) %>% summarise('successful pregnancies' = n()) %>% pull('successful pregnancies')

com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 1 & LMWH == 1) %>% filter(live == 0) %>% summarise('failed pregnancies' = n())

com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 1 & LMWH == 1) %>% summarise('success rate' = mean(live)*100)

com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 1 & LMWH == 1) %>% filter(live == 0 & POA < 10) %>% summarise('MC before 10 wk' = n())

com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 1 & LMWH == 1) %>% filter(live == 0 & POA >= 10) %>% summarise('MC at or after 10 wk' = n())
```


Previous pregnancy outcome - Treated Aspirin alone
```{r}
com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 1 & LMWH == 0) %>% summarise('total pregnancies' = n())

com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 1 & LMWH == 0) %>% filter(live == 1) %>% summarise('successful pregnancies' = n())

asp_treatment_total <- com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 1 & LMWH == 0) %>% summarise('total pregnancies' = n()) %>% pull(`total pregnancies`)

asp_treatment_success <- com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 1 & LMWH == 0) %>% filter(live == 1) %>% summarise('successful pregnancies' = n()) %>% pull('successful pregnancies')

com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 1 & LMWH == 0) %>% filter(live == 0) %>% summarise('failed pregnancies' = n())

com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 1 & LMWH == 0) %>% summarise('success rate' = mean(live)*100)

com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 1 & LMWH == 0) %>% filter(live == 0 & POA < 10) %>% summarise('MC before 10 wk' = n())

com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 1 & LMWH == 0) %>% filter(live == 0 & POA >= 10) %>% summarise('MC at or after 10 wk' = n())
```


Previous pregnancy outcome - Treated LMWH alone
```{r}
com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 0 & LMWH == 1) %>% summarise('total pregnancies' = n())

com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 0 & LMWH == 1) %>% filter(live == 1) %>% summarise('successful pregnancies' = n())


LMWH_treatment_total <- com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 0 & LMWH == 1) %>% summarise('total pregnancies' = n()) %>% pull(`total pregnancies`)

LMWH_treatment_success <- com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 0 & LMWH == 1) %>% filter(live == 1) %>% summarise('successful pregnancies' = n()) %>% pull('successful pregnancies')

com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 0 & LMWH == 1) %>% filter(live == 0) %>% summarise('failed pregnancies' = n())

com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 0 & LMWH == 1) %>% summarise('success rate' = mean(live)*100)

com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 0 & LMWH == 1) %>% filter(live == 0 & POA < 10) %>% summarise('MC before 10 wk' = n())

com %>% mutate(ASP = ifelse(is.na(ASP),0,ASP), LMWH = ifelse(is.na(LMWH),0,LMWH)) %>% filter(ASP == 0 & LMWH == 1) %>% filter(live == 0 & POA >= 10) %>% summarise('MC at or after 10 wk' = n())
```


comparison

No treatment vs any
```{r}
f1 <- prop.test(c(either_treatment_success, no_treatment_success), c(either_treatment_total, no_treatment_total), p = NULL, alternative = "greater",correct = TRUE)
print(tidy(f1))
```

No treatment vs both
```{r}
f1 <- prop.test(c(both_treatment_success, no_treatment_success), c(both_treatment_total, no_treatment_total), p = NULL, alternative = "greater",correct = TRUE)
print(tidy(f1))
```

No treatment vs asp
```{r}
f1 <- prop.test(c(asp_treatment_success, no_treatment_success), c(asp_treatment_total, no_treatment_total), p = NULL, alternative = "greater",correct = TRUE)
print(tidy(f1))
```


No treatment vs LMWH
```{r}
f1 <- prop.test(c(LMWH_treatment_success, no_treatment_success), c(LMWH_treatment_total, no_treatment_total), p = NULL, alternative = "greater",correct = TRUE)
print(tidy(f1))
```

ASP treatment vs LMWH
```{r}
f1 <- prop.test(c(LMWH_treatment_success, asp_treatment_success), c(LMWH_treatment_total, asp_treatment_total), p = NULL, alternative = "two.sided",correct = TRUE)
print(tidy(f1))
```

ASP treatment vs both
```{r}
f1 <- prop.test(c(both_treatment_success, asp_treatment_success), c(both_treatment_total, asp_treatment_total), p = NULL, alternative = "greater",correct = TRUE)
print(tidy(f1))
```

LMWH treatment vs both
```{r}
f1 <- prop.test(c(both_treatment_success, LMWH_treatment_success), c(both_treatment_total, LMWH_treatment_total), p = NULL, alternative = "greater",correct = TRUE)
print(tidy(f1))
```



\ \

#### results
A total of `r total_pregnancies` pregnancies were evaluated of `r total_mothers` mothers. The median number of pregnancies was `r pregnancy_quantiles["50%"]` with an inter-quantile range of 2.
```{r fig.height=3, fig.width=3}
pregnancy_boxplot
```
The mean age of the selected mothers was `r age$mean` with a standard deviation of `r age$sd`.
```{r fig.height=3, fig.width=5}
age_histogram + labs(x = "Age")
```




`r treatment$Both` pregnancies were treated with both LMWH and aspirin. `r treatment$Aspirin` and `r treatment$LMWH` pregnancies were treated with aspirin alone and LMWH alone, respectively.

\ 
`r pih$PIH` pregnancies were complicated with PIH.
```{r}
pih
```
\
Outcome by treatment group
```{r fig.height=5, fig.width=5}
success <- c(asp_treatment_success, LMWH_treatment_success, either_treatment_success, no_treatment_success, both_treatment_success)

treatment_table <- treatment %>%  pivot_longer(cols = names(treatment), names_to = 'Category', values_to = 'Total') %>% filter(! Category %in% c('Aspirin', 'LMWH')) %>% cbind(success) %>% mutate(rate = success * 100 / Total)

treatment_table %>%
  filter(Category != 'Either') %>% 
  mutate(failed = Total - success) %>%
  pivot_longer(cols = c('success','failed'), names_to = 'state', values_to = "number") %>%
  ggplot(aes(x = reorder(Category, rate), y = number, fill = factor(state, levels=c('success', 'failed'))))  +
  geom_bar(stat = 'identity') + 
  labs(x = 'Category', fill = "")
```

The rate of live births among mothers who were not received specific treatment for APS was `r no_treatment_success / no_treatment_total`. In contrast, `r either_treatment_total` gestations were treated with either aspirin, LMWH or both and the rate of successful outcome was `r either_treatment_success / either_treatment_total`

Outcome in each group is summarized below

```{r}
knitr::kable(treatment_table %>% mutate(Category = str_replace_all(.$Category, "_", " ")) %>% rename(n = Total, 'live births' = success, 'live birth rate' = rate) %>% arrange(`live birth rate`) %>% filter(Category != 'Either'))
```

\
Investigations
\
`r pull(investigations_all["Any investigation out of Lupus, cardio, beta, APLS"]) ` (`r pull(investigations_all["precent - of mothers"])`%) mothers had undergone any of antibody tests for APL and out of them, `r pull(investigations_positive["positive subjects"])` (`r pull(investigations_positive["percent of tested"])`%) yeilded a positive result.

```{r}
knitr::kable(full %>% summarise(Lupus = sum(!is.na(Lupus)), cardiolipin = sum(!is.na(Cardiolipin)), 'beta 2 microglobulin' = sum(!is.na(`beta2 microglobulin`)), APL = sum(!is.na(APLS)) ))
```

\

Live rates between test positives and negatives

```{r}
p <- com %>%
  mutate(any = ifelse(!is.na(Lupus) | !is.na((Cardiolipin)) | !is.na((APLS)) | !is.na((`beta2 microglobulin`)), 1,0 ), any_positive = ifelse((Lupus) == 1 | (Cardiolipin) == 1  | (APLS) == 1 | (`beta2 microglobulin`) == 1, 1, 0)) %>%
  mutate(any_positive = ifelse(is.na(any_positive),0, any_positive)) %>%
  filter(any == 1) %>% filter(ASP == 1 | LMWH == 1) %>% group_by(any_positive) %>% summarise(n = n(), live_number = sum(live), live_rate = mean(live))
p1 <- (tidy(prop.test(p$live_number, p$n)))
```

Number of pregnancies that were treated, in mothers who were tested for antibodies was `r sum(p$n)`.
Out of them, `r p$n[2]` (`r p$n[2] * 100 / sum(p$n)`%) mothers tested positive for at least one test and the success rate was `r p$live_rate[2]`.
`r p$n[1]` (`r p$n[1] * 100 / sum(p$n)`%) mothers had negative results and success rate among them was `r p$live_rate[1]`

\ \ 

